---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
```{r, echo=FALSE}
library(jsonlite)
library(stringr)
library(glmnet)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(nnet)
```


**Loading data sets**

```{r, echo=FALSE}
user_data <- load("yelp_user_small.Rda")
review_data <- load("yelp_review_small.Rda")
business_data <- stream_in(file("yelp_academic_dataset_business.json")) 
```

**Checking for duplicate entries**

```{r}
busdup <- sum(duplicated(business_data$business_id))
revdup <- sum(duplicated(review_data_small$business_id))
usdup <- sum(duplicated(review_data_small$user_id))
```

business_data: 0 business have been included twice
review_data: \( \frac{1,261,472}{1,398,056} \) have been reviewed more than once. \( \frac{686,446}{1,398,056} \) users have left more than 1 review

```{r}
sum(is.na(user_data_small))
sum(is.na(review_data_small))
```
*there are no missing values in user_data or review_data


user_data_small$elite has only 18,178 entries for "elite" column out of 397,579, although not coded as "na" they can be viewed as missing values and can be dropped
```{r}
user_data_small$elite <- NULL
```

user_data_small$friends can be transformed to display number of friends although the attributes of a user's friends likely has predictive power it will be difficult to model

coding all "none" values as 0, and adding a friend count
```{r}
user_data_small$friends <- ifelse(user_data_small$friends=="None",0,str_count(user_data_small$friends, ",") + 1)
```

Selecting relevant predictors from business_data dataset
```{r}
business_data_new <- subset(business_data, select = c("city", "stars", "review_count", "is_open", "business_id"))
```

merging review data set with user data set and business data set

```{r}
merged_df <- merge(merge(review_data_small, user_data_small, by = "user_id"), business_data_new, by = "business_id")

# making "city" into a factor variable
merged_df$city <- factor(merged_df$city)
```

counting the number of zero values in each column

```{r}
# Count the number of zeros in each column
num_zeros_per_column <- round((colSums(merged_df == 0)/nrow(merged_df))*100,2)

x_values = names(num_zeros_per_column)
y_values = unname(num_zeros_per_column)

col_sums <- data.frame(variable = x_values, proportion = y_values)
col_sums <- col_sums[col_sums$proportion != 0,]


# creating a plot

zeros_plot <- ggplot(data = col_sums, mapping = aes(x=reorder(variable,-proportion), y = proportion)) + geom_bar(stat = "identity", fill = "skyblue") + geom_hline(yintercept = 30, linetype="dashed", color="red", size =1) + 
  ggtitle("Percentage of 0 values per column") + 
  labs(x="Variable name", y = "Percentage (%)") +
  theme(axis.text.x=element_text(angle=45, hjust=1), 
        plot.title    = element_text(family = "mono", hjust =0.5),
        axis.title.x = element_text(family="mono"),
        axis.title.y = element_text(family="mono")
        ) 

#ggsave("zeros_plot.png", plot = zeros_plot, width = 4, height = 4)


zeros_plot
```
Removing irrelevant variables
dropping "city£ because certain cities will not be common to both train and test data
and will generate an error

```{r}
clean_df <- merged_df[ ,(colSums(merged_df == 0)/nrow(merged_df) *100) < 30] %>% select(-c("business_id","user_id", "review_id", "text", "date", "name", "yelping_since", "city")) %>%
  rename(review_stars = stars.x, review_count_user = review_count.x, useful = useful.y, funny = funny.y, cool = cool.y , user_stars = average_stars, review_count_business = review_count.y, business_stars = stars.y)
```



splitting into test and train

```{r}
set.seed(1)
train_split <- sample(1:nrow(clean_df), 3*nrow(clean_df)/4)

# dropping irrelevant variables
train <- clean_df[train_split,] 
test <- clean_df[(-train_split),] 

trainX <- subset(train,select = -review_stars)
trainY <- subset(train, select = review_stars)
testX <- subset(test, select = -review_stars)
testY <- subset(test, select = review_stars)
```

Random Choice model for comparison

```{r}
randY <- sample(1:5, length(testY$review_stars), replace = TRUE)
```

random choice MSE

```{r}
rand_MSE<-mean((randY-testY$review_stars)^2)

rand_MSE
```


fitting multinomial logit model due to discreet outcome variables

```{r}
ml_model <- multinom(review_stars ~., data=train)
```

fitting a linear model

```{r}
linmod <- lm(review_stars ~., data=train)
```

LM predictions

```{r}
linmod_predict <- predict(linmod, newdata = testX)
```

LM MSE calculations

```{r}
test_MSE<-mean((linmod_predict-testY$review_stars)^2)

test_MSE
```
multinom predictions

```{r}
mlmod_predict <- predict(ml_model, newdata = testX)
```

